---
title: "Partisan vs Climate Cues and Public Opinion about Trump's Actions on Energy"
bibliography: references.bib
csl: american-political-science-association.csl
fig-cap-location: top
fontfamily: ebgaramond
link-bibliography: false
format:
  html:
    author:
      - name: Matthew C. Nowlin
        email: matthew.nowlin@uta.edu
        corresponding: true
        affiliations:
          - name: University of Texas at Arlington
            department: Department of Political Science
  pdf:
    author:
      - name: "Matthew C. Nowlin^[Associate Professor, Department of Political Science, University of Texas at Arlington. Email: matthew.nowlin@uta.edu]"
    include-in-header:
      text: |
        \usepackage{setspace}
        \usepackage{dcolumn}
        \usepackage{float}
        \usepackage{amsmath}
        \addtokomafont{disposition}{\rmfamily}
        \normalem
  docx:
    author:
      - name: "Matthew C. Nowlin^[Associate Professor, Department of Political Science, University of Texas at Arlington. Email: matthew.nowlin@uta.edu]"
    reference-doc: custom-reference-doc.docx
---

```{r}
#| include: false
source("scripts/manuscript-prep.R")
```

**Abstract**: The second Trump administration has made strong and concerted efforts to shape energy production in the United States. In particular, the administration has moved to strengthen the fossil fuel industry and limit the use of renewable energy sources. Increasing the use of renewable sources is needed to address climate change. Yet, even with the connection to climate change, public support for renewable energy and policies supporting the development of renewable energy sources is not as polarized as support for climate change policies. However, the efforts of the Trump administration may serve to polarize public opinion on energy sources and the government's role in shaping their availability. In this paper, I develop a preregistered survey experiment to examine the polarization in support for several recent administration actions on energy when respondents are given a Trump cue, a climate change cue, and no cues. Specifically, I surveyed 3,000 respondents recruited from Cloud Research, with 1,000 participants receiving a Trump cue, 1,000 participants receiving a climate cue, and 1,000 respondents in the control condition, who received no cue. Respondents in each condition were asked about their level of support for several actions, including limiting tax credits for wind and solar projects, requiring coal plants to remain open, opening federal lands for coal leasing, cancelling a nearly complete offshore wind project, and shortening the timeline of regulatory approval for new nuclear plants. I compare the impacts of the Trump cue and climate cue on support for Trump administration actions across partisanship, ideology, and education level. I find that the Trump cue, but not the climate cue, increased support for Trump's energy actions among conservative Republicans, most notably for opening federal lands to coal leasing and for shortening nuclear licensing timelines. The climate cue had little effect on Republicans, except for a decrease in support for eliminating tax credits for wind and solar projects, the action that most directly limits renewable energy. Liberal Democrats were largely unresponsive to either cue, and neither cue's effects were conditioned by whether respondents held a college degree. These results suggest that Trump's actions on energy are more likely to polarize public opinion through their association with Trump himself than through their connection to climate change, and that this pattern holds regardless of educational attainment.

# Introduction 

Since returning to office, President Trump has sought to promote fossil fuels while sidelining renewable sources of energy. He has also downplayed the threat of climate change and through executive actions has worked to limit approaches to address climate change across the federal government. For example, early in his second term, Trump declared a national energy emergency, directing federal agencies to use emergency statutory authorities to speed the development and authorization of energy projects and rescinding a number of prior climate-related executive orders [@houseDeclaringNationalEnergy2025]. Historically, energy has not been polarized as public opinion on energy tended to be focused on cost and availability [@ansolabehereCheapCleanHow2016]. Yet, in recent years energy sources have become more polarized with Republicans tending to favor fossil fuel sources, and Democrats favoring renewable sources. This increasing polarization may be a result of the connection between renewable energy and climate change. Additionally, Trump's actions in his second term may serve to further polarize the public on energy issues. 

The potential for Trump's actions to polarize comes from the fact that he himself is such a polarizing figure. As people connect those actions to Trump himself, they are cued to support/oppose them based on their views of Trump. Previous research has noted the importance of cues in shaping public opinion. Cues provide a shortcut which allows the mass public, who are likely ignorant of the details of politics and policy issues, to infer reasonably consistent views on a host of issues. Partisan cues in particular are known to be important sources of information for the public [@kamWhoToesParty2005; @schaffnerPartisanHeuristicLowInformation2002]. Additionally, the impact of cues may vary by level of education and political sophistication, with those less knowledgeable relying more heavily on cues than those with more knowledge [@kamWhoToesParty2005; @merolaNumeracyPersuasiveEffect2016]. With Trump's support among those without a college education, it is possible that a Trump cue will be particularly influential on that group [@morganTrumpVotersWhite2018; @zingherTRENDSDiplomaDivide2022].   

Given the polarized views of Trump and climate change, in this paper, I examine the role that a Trump and a climate change cue have on support for several of Trump's actions on energy in the first year of his second term. The actions I examine include limiting tax credits for wind and solar projects, requiring coal plants to remain open, opening federal lands for coal leasing, cancelling a nearly complete offshore wind project, and shortening the timeline of regulatory approval for new nuclear plants. Specifically, I use a survey experiment where respondents were randomly assigned to receive a Trump cue (a statement that the actions came from the Trump Administration versus the federal government), a climate cue (a statement about climate scientists being concerned that some of the actions could make climate change worse), or to a control condition with no cue. I test several preregistered hypotheses regarding how respondents' support would shift, compared to those in the control group, in response to the Trump and climate cues based on political beliefs and education.  

Overall, I find that the Trump cue, but not the climate cue, increased support for Trump's energy actions among conservative Republicans, most notably for opening federal lands to coal leasing and for shortening nuclear licensing timelines. The climate cue had little effect on Republicans, except for a _decrease_ in support for eliminating tax credits for wind and solar projects, which is the action that most directly limits renewable energy. Liberal Democrats were largely unresponsive to either cue. Finally, neither cue's effects were conditioned by whether respondents held a college degree. These results suggest that Trump's actions on energy are more likely to polarize public opinion through their association with Trump himself than through their connection to climate change. Additionally, that impact is consistent between those with and without a college degree. 

# Polarization on Energy Issues

Public views on energy in the US have grown increasingly polarized along partisan lines, even as aggregate support for many energy sources and policies has remained relatively stable. Using data from three long-running national surveys spanning nearly two decades, Bergquist, Konisky, and Kotcher [-@bergquistEnergyPolicyPublic2020] found that, although there was overall public support for shifting from fossil fuels to renewable sources of energy, there were differences in support between Republicans and Democrats, and that difference had grown in the last decade. Additionally, they show that when the costs of clean energy policies are made explicit, support declines by a similar magnitude for both Democrats and Republicans, suggesting that the partisan divide in energy opinion is not simply a function of differential cost sensitivity [also see @benegalCostSensitivityPartisan2024].

Yet this partisan divide is not the same across every energy source [@ansolabehereCheapCleanHow2016; @bergquistEnergyPolicyPublic2020]. Fossil fuels carry the deepest partisan split of any energy source. For example, Democrats and liberals are less supportive of coal than Republicans and conservatives [@hartPoliticalPolarizationSupport2021], and are more concerned about the extraction methods behind oil and gas development [@davisEnergyAbundanceEnvironmental2014; @raimiPartisanshipProximityPredict2020]. Underlying this gap is a difference in what is salient with regard to energy sources. For Democrats and liberals the health, environmental, and climate harms of fossil fuels, are the most salient considerations while Republicans and conservatives point to the jobs, prices, and energy security they provide [@clarkePoliticsScientificConsensus2019; @howellSeeingRiskColoredGlasses2019]. With regard to wind and solar, majorities across the partisan divide have favored expanding these sources, and where partisans disagree, it is a matter of how much to expand renewables rather than whether to do so at all [@bergquistEnergyPolicyPublic2020; @hamiltonChangeWindUS2018; @hawesClimateSciencePolitics2022]. Nuclear power falls in between, drawing modestly more support from the right than the left, but without the sharp partisan split seen for fossil fuels [@yeoPartisanAmplificationRisk2014; @mcbethPoliticalIdeologyNuclear2023a].

Underlying this partisan divide are different reasons for supporting, or opposing, a transition toward renewable energy. As noted by Gustafson et al. [-@gustafsonRepublicansDemocratsDiffer2020], Republicans' support for renewable energy is driven more by considerations of economic costs and benefits, whereas Democrats' support is driven more by concern about climate change. Using a nationally representative survey of registered voters, they found that Democrats most often cite "reduce global warming" as their single most important reason to support renewable energy (chosen by 33 percent of Democrats, compared to 7 percent of Republicans), while Republicans more often cite reasons tied to energy costs and energy independence. This divergence in underlying motivation is likely important for understanding how partisan and climate cues will separately shape support for energy actions. For example, a cue based on cost or energy security is likely to resonate differently with Republicans than a climate change cue.

Other work argues that this partisan divide in energy attitudes reflects several distinct dimensions of partisanship rather than a single underlying mechanism. Specifically, Mayer [-@mayerPartisanshipPoliticsEnergy2019] proposes a four-part conceptual framework encompassing cue receptivity, expressive partisanship, negative partisanship, and performative partisanship, each of which may separately shape how citizens respond to information about energy. This framework is useful for understanding how the cues examined here, namely a Trump cue and a climate change cue, are likely to operate through different channels. The Trump cue is likely to activate expressive and cue-receptive partisanship among Republicans, while a climate cue is likely to activate the substantive climate concern among Democrats.

The unevenness in baseline polarization on energy will likely impact the ways in which the cues work on support for Trump's actions. Two of the actions, keeping coal plants open and opening federal land to coal leasing, support fossil fuels, the most polarized energy sources. Therefore, partisan reactions to a Trump or climate cue will likely be sharpest. Two others, limiting wind and solar tax credits and cancelling the offshore wind project, limit renewable sources where the public, Republicans included, has tended to be supportive [@bergquistEnergyPolicyPublic2020; @hamiltonChangeWindUS2018]. As a result, the cue effects there may be more muted for those actions. The fifth action, accelerating nuclear licensing, impacts the least polarized energy source [@yeoPartisanAmplificationRisk2014; @mcbethPoliticalIdeologyNuclear2023a], so the expectations are less clear.

In the next section, I discuss the theoretical mechanisms of partisan cue-taking in greater detail and describe how the Trump and climate cues are likely to shape support for Trump's actions on energy in different ways.

# Cues and Cue-Taking

Partisan cues function as a heuristic that allows the mass public to form opinions on complex policy issues without extensive information [@kamWhoToesParty2005; @schaffnerPartisanHeuristicLowInformation2002]. Rather than evaluating the substantive merits of an issue, citizens often infer their own position from the position taken by their preferred party or its elites [@slothuusPoliticalPartiesMotivated2010; @leeperPoliticalPartiesMotivated2014]. This cue-taking is best understood through the lens of motivated reasoning, wherein citizens process information in a way that protects and reinforces their prior partisan commitments rather than arriving at accuracy [@taberMotivatedSkepticismEvaluation2006; @petersenMotivatedReasoningPolitical2013]. As elite polarization has intensified, party cues have become an increasingly powerful, and at times dominant, force in opinion formation, often overwhelming the effect of substantive policy information [@druckmanHowElitePartisan2013; @cohenPartyPolicyDominating2003].

Climate change and energy policy are particularly susceptible to this dynamic. As shown by Feldman and Hart [-@feldmanClimateChangePolarizing2018], attaching a climate change frame or a cue from a Democratic source to an energy policy is often enough to depress support among Republicans, even when the underlying policy is left unchanged. Similarly, out-group cues from Democratic elites have been found to produce a durable backlash in climate skepticism among Republicans [@merkleyPartyCuesNews2021]. Additionally, policy proposals can become polarized as they become associated with one political party, such as the Green New Deal [@gustafsonDevelopmentPartisanPolarization2019; @mcconnellGreenNewDeal2022]. These findings suggest that Trump cues and climate cues are likely to activate partisan reactions to actions on energy. Specifically, conservative Republicans are likely to become more supportive when an action is cued as coming from Trump or as countering climate change, whereas liberal Democrats are likely to become less supportive.

A cue naming Trump specifically is theoretically distinct from a generic partisan cue. Individual elites, rather than parties in the abstract, have been shown to carry more weight in shaping public opinion [@nicholsonPolarizingCues2012], and elite cues more broadly rank among the strongest known predictors of public views on climate change [@brulleShiftingPublicOpinion2012]. Trump in particular has been shown to move opinion on issues, including immigration [@taylorTrumpEffectImmigration2026] and trade policy [@essigTrumpEffectPolitical2021]. In addition, presidents more generally appear able to shape partisan divides on climate policy simply by becoming associated with it [@childreePolarizingFiguresPolarized2026]. Given that his administration has taken concrete steps to expand fossil fuel production while curbing renewable energy development [@gerrardTrumpsMultiProngedAttack2025; @joselowHowTrumpsFirst2025], it is likely that this same dynamic will carry over into energy attitudes specifically.

Some early accounts of the partisan heuristic emphasized that cue-taking substitutes for the cognitive effort required to process substantive information, implying that citizens with less knowledge about politics, or less education, should rely on party cues the most [@kamWhoToesParty2005]. Research using numeracy also supports this account, as low-numeracy respondents rely more heavily on party cues, while high-numeracy respondents are more likely to update toward substantive policy information when it is available [@merolaNumeracyPersuasiveEffect2016]. Yet effortful thinking alone does not appear to be sufficient to reduce cue reliance; rather, it may be the availability of policy information, not cognitive effort per se, that weakens party cue effects [@tappinEstimatingCausalEffects2026]. Additionally, political sophistication and education are also associated with more effective motivated reasoning, meaning that better-informed citizens may process supportive information more efficiently rather than less [@kahanIdeologyMotivatedReasoning2013]. Bakker, Lelkes, and Malka [-@bakkerUnderstandingPartisanCue2019] similarly find that cue receptivity is not simply a function of bounded rationality but is strongest when cognitive resources are placed in service of expressive, identity-affirming reasoning. Consistent with this more conditional pattern, education has been found to amplify, rather than dampen, partisan polarization over climate change, particularly among conservatives [@hamiltonEducationPoliticsOpinions2011]. This amplifying pattern is not limited to climate change specifically as Ehret, Sparks, and Sherman [-@ehretSupportEnvironmentalProtection2017] find that more educated liberals and conservatives diverge further from one another on environmental attitudes generally, consistent with an ideological-consistency model in which education strengthens attentiveness to elite cues rather than substituting for them. Yet education's moderating role may itself be conditional on whether a cue is congruent with a respondent's existing predispositions. Vrânceanu [-@vranceanuHowCrosspressuresAffect2022] finds that less-educated partisans shift the most when exposed to a cue that conflicts with the values typically associated with their level of education, while highly educated partisans are comparatively insensitive to such incongruent cues. Together, this literature suggests that a college degree is likely to shape responsiveness to Trump and climate cues, but the direction of that relationship is less straightforward than what is implied by the low-information heuristic model.

A Trump cue may act differently than a typical partisan cue because of Trump's purported strength among "working-class" voters, particularly white working-class voters. This pattern is consistent with a broader realignment of the American electorate along educational lines [@grossmannPolarizedDegreesHow2024]. Kitschelt and Rehm [-@kitscheltSecularPartisanRealignment2019] show that education, particularly when combined with income, has become one of the strongest predictors of white partisan support since the New Deal era, with non-college whites increasingly aligned with the Republican Party. Zingher [-@zingherTRENDSDiplomaDivide2022] similarly documents a widening "diploma divide," finding that educational attainment has become an increasingly strong predictor of both partisanship and vote choice since 2000, with less-educated counties trending Republican and more-educated counties trending Democratic. Sances [-@sancesHowUnusualWas2019] finds that the education-party correlation was the strongest on record in the 2016 election, and Morgan and Lee [-@morganTrumpVotersWhite2018] show that white working-class voters, including many who had supported Obama in 2012, were disproportionately likely to switch to Trump.  

If support for Trump reflects an education divide over in addition to a straightforward partisan signal, a Trump cue may activate identity-based responses among non-college conservative Republicans and liberal Democrats that a generic partisan or climate cue would not. Whether this education-based moderation is actually present, however, is an empirical question that has not, to my knowledge, been tested directly in the context of cue-taking on energy policy.

## Hypotheses

Taken together, the literature points to two ways through which the Trump and climate cues could shape support for Trump's actions on energy. The first is straightforward partisan cue-taking, as a Trump cue is likely to activate expressive and cue-receptive partisanship among Republicans while signaling an out-group threat to Democrats. A climate cue is likely to tap the substantive climate concern that already distinguishes Democrats from Republicans. Therefore, conservative Republicans and liberal Democrats should respond most strongly to both cues, and in opposite directions. The second is education tied to the role of cues for those with less education as well as Trump's support among those without a college degree. If a Trump cue draws on an identity-based appeal among non-college conservatives rather than functioning as an ordinary partisan signal, its effects should be concentrated among respondents without a college degree. I test both the partisan and education explanations with the following preregistered hypotheses.^[The "than others" in the hypotheses refer to those with other political beliefs, educational attainment, or both.] <!-- ^[See https://osf.io/zkhqp/overview] --> 

> H1: Conservative Republicans who receive the Trump or climate change cue will be more supportive of Trump's actions on energy than others 

> H2: Liberal Democrats who receive the Trump or climate change cue will be less supportive of Trump's actions on energy than others 

> H3: Conservative Republicans without a college degree who receive the Trump cue will be more supportive of Trump's actions on energy than others

> H4: Liberal Democrats without a college degree who receive the Trump cue will be less supportive of Trump's actions on energy than others 


# Data and Measures 

The data for this study was obtained from a survey of a quota-based sample of US residents recruited through Cloud Research. The quota-based sampling approach used US Census data to help ensure representation across key demographic groups, including age, gender, and race. The survey was internet-based and administered using QuestionPro and was fielded from April 10 to April 15, 2026. Survey weights were applied to ensure that the sample was representative of the US population. Specifically, I used the `survey` and `anesrake` packages in `R` to perform iterative raking to compute survey weights that align the sample with the US Census's 2023 American Community Survey (ACS) on key demographic variables, including age, race, gender, and education. Additionally, I used the Pew Research Center's National Public Opinion Reference Survey (NPORS) from 2025 to align the sample on political party affiliation. 

To measure the impact of cues on support for Trump Administration actions on energy, respondents were randomly assigned to one of three conditions, including a Trump cue, a climate cue, and control condition with no cue. Overall, the groups were roughly equal, with the Trump cue, *n* = `r n_trump`, the climate cue, *n* = `r n_climate`, and the control, *n* = `r n_control`, from the sample of *N* = `r n_total`. The Trump cue identified the actions as being done by the _Trump Administration_, whereas the climate cue and the control conditions indicated the actions were by the _federal government_. The climate cue included a statement about concerns that climate scientists have about the impact of the actions on climate change. 

The actions came from a number of well-publicized steps taken by the Trump Administration in the first year of his second term. These included actions by the Department of the Treasury that tightened the rules that wind and solar developers rely on to qualify for federal tax credits [@springTrumpAdministrationLimits2025], the Department of Energy invoking emergency authority to keep aging coal plants running past their planned retirement dates, including plants in Michigan, Washington, Indiana, and Colorado [@thakoreTrumpAdministrationOrders2025]. The Department of the Interior opened 13.1 million acres of federal land in Montana, North Dakota, and Wyoming for coal leasing [@InteriorUnleashesAmerican2025], and separately halted construction on offshore wind projects along the East Coast, two of which were nearly complete [@wasserTrumpAdministrationPauses2025]. At the same time, the administration sought to accelerate nuclear energy development, directing the Nuclear Regulatory Commission to shorten its licensing review to under 18 months for new reactors [@usdepartmentofenergy9KeyTakeaways2025; @TrumpsLicensingGoals2025].

The question wording for support of these actions was as follows: 

> Recently, the [_Trump Administration_/_Federal government_] took several actions to shape electricity generation in the U.S.

> **Climate cue**: _Scientists are concerned that the actions associated with coal, wind, and solar energy will make climate change worse._

> Using a scale of 1 to 5, where 1 means strongly oppose and 5 means strongly support, how do you feel about the following actions of the [_Trump Administration_/_Federal government_]? [The following were randomized]

> * The [_Trump Administration_/_Federal government_] limiting the eligibility of wind and solar energy projects for federal tax credits
> * The [_Trump Administration_/_Federal government_] requiring coal plants that were set to retire to remain open
> * The [_Trump Administration_/_Federal government_] opening 13.1 million acres of federal land for coal leasing
> * The [_Trump Administration_/_Federal government_] cancelling an offshore wind project that was nearly complete
> * The [_Trump Administration_/_Federal government_] shortening the time it takes to approve and license the construction of new nuclear power plants

The main independent variables of interest are political beliefs, including conservative Republicans and liberal Democrats, with other political beliefs (moderate independents, conservative Democrats, liberal Republicans) as the excluded referents. The political belief measures were constructed based on responses to a self-reported ideology scale with 1 = strongly liberal, 4 = middle of the road, and 7 = strongly conservative and with a question asking with which political party they most identify. College is a dummy variable, 1 = completed a college degree, 0 = did not. @tbl-descriptives shows the descriptive statistics for each of the variables. 

```{r}
#| label: tbl-descriptives
#| tbl-cap: "Descriptive Statistics for Dependent and Independent Variables"
desc_table
```

In terms of support for the various actions, they each seem to hover around the mid-point on the 1 to 5 scale with cancelling nearly complete wind projects showing the least support at `r round(wtd_mean(d$fed.action.cancel.wind), 2)` and reducing nuclear energy regulations showing the most at `r round(wtd_mean(d$fed.action.nuclear.licensing), 2)`.

# Results

I first examine the difference in means across the cue conditions, and the results are shown in @tbl-means. As can be seen, the only significant difference was in support for the nuclear regulatory changes, which was lower in the Trump condition compared to the control. 

```{r}
#| label: tbl-means
#| tbl-cap: "Weighted Mean Support for Trump's Energy Actions by Cue Condition"
means_table
```

The first two hypotheses, H1 and H2, posited that political beliefs will shape the impact of the Trump and climate cues, with conservative Republicans expected to respond to the Trump and climate cues with more support for Trump actions and liberal Democrats expected to respond with less support. @tbl-results shows the full results for each action, and @fig-identity shows the results of the OLS models for cue condition and political beliefs.

```{r}
#| label: tbl-results
#| tbl-cap: "Weighted OLS Models of Support for Trump's Energy Actions"
results_table
```

As shown in @tbl-results, the main effect of the Trump cue was negative and significant (at _p_<0.10) for reducing nuclear regulations. For the main effect of the climate cue, it was associated with an increase in support for limiting tax credits for wind and solar and for keeping coal plants open (at _p_<0.10), all else equal. Additionally, the main effect for political beliefs was that conservative Republicans were more likely than those with moderate political beliefs to support all of Trump's actions, whereas liberal Democrats were less likely to support his actions. Next, I discuss the interactions between political beliefs and cues, which are illustrated in @fig-identity.  

```{r}
#| label: fig-identity
#| fig-cap: "Predicted Support by Cue Condition and Political Beliefs"
#| fig-width: 11
#| fig-height: 4
fig_identity
```

For conservative Republicans, the Trump cue significantly increased their support for opening federal lands to coal leasing and for shortening approval for nuclear power construction. Additionally, support for keeping coal plants open and cancelling the offshore wind project was significant at _p_<0.10 among conservative Republicans in the Trump condition. However, the Trump cue did not influence support for eliminating federal tax credits; rather, the climate cue led to a decrease in support for eliminating the tax credits among conservative Republicans. Otherwise, the climate cue did not have any effects for Republicans, contrary to expectations.  

Liberal Democrats were largely unresponsive to either the Trump cue or the climate cue, compared to those in the control condition. However, the climate cue reduced support for eliminating the tax credits at _p_<0.10. Next, I examine the hypotheses, H3 and H4, regarding the effect of education and cues. The results are shown in @fig-college. 

```{r}
#| label: fig-college
#| fig-cap: "Predicted Support by Trump Cue and College Degree, Conservative Republicans and Liberal Democrats"
#| fig-width: 11
#| fig-height: 6
fig_college
```

As can be seen, there were no differences in support across the three-way interaction of the Trump cue, college education, and political beliefs. This means that those without a college degree were no more influenced by the Trump cue than those with a degree. This finding held for Republicans and Democrats.^[I also examined the interaction of education and the climate cue. Like the Trump cue, there were no significant differences. The results are shown in the Appendix, @tbl-results-explore.] 

While I had no prior expectations regarding the two-way interaction between education and political beliefs, there are some significant effects. As shown in @tbl-results, college educated conservative Republicans were less supportive of cancelling nearly complete wind projects and nuclear regulation changes (at _p_<0.10) than non-college educated Republicans. Additionally, college educated liberal Democrats were less supportive of limiting tax credits, cancelling wind projects, and nuclear regulations.    

Overall, the evidence supported the Trump cue aspect of H1, as the Trump cue increased support among conservative Republicans for four of the five actions, including coal leasing, keeping coal plants open, cancelling the offshore wind project, and nuclear licensing. However, there was no support for the other hypotheses, as neither cue impacted liberal Democrats' support (H2), nor did education moderate the Trump cue's effect on either group (H3, H4). 


# Discussion and Conclusion

In this paper, I examined whether a Trump cue or a climate change cue shapes public support for five of the second Trump administration's actions on energy, and whether these effects vary by political beliefs and education. Overall, the clearest pattern in the results is that the Trump cue, but not the climate cue, increased support for Trump's actions among conservative Republicans, while liberal Democrats and less-educated respondents were largely unresponsive to either cue.

Consistent with H1, conservative Republicans exposed to the Trump cue were significantly more supportive of opening federal lands to coal leasing and of shortening the regulatory timeline for nuclear power construction, with marginally significant increases in support for keeping coal plants open and cancelling the offshore wind project. This finding is consistent with the expectation that a cue naming Trump specifically activates expressive and cue-receptive partisanship among Republicans [@mayerPartisanshipPoliticsEnergy2019], and it fits the broader elite cues literature showing that cues from individual political figures, rather than generic partisan cues, are especially likely to move opinion [@nicholsonPolarizingCues2012]. Notably, however, the Trump cue did not increase Republican support for eliminating tax credits for wind and solar projects. One likely explanation is that eliminating tax credits is the only one of the five actions that directly reduces, rather than expands or maintains, an existing energy program, and this difference in the nature of the action may be an important factor for respondents.

The climate cue produced no significant effects among Republicans for four of the five actions, contrary to expectations. However, it significantly _decreased_ Republican support for eliminating tax credits, the one action that most directly harms renewable energy deployment. This could be because the climate cue explicitly named coal, wind, and solar energy and warned that the administration's actions could make climate change worse; for conservative Republicans, this framing may have been credible specifically for an action that curtails renewables, even as it failed to move opinion on the actions involving coal or nuclear. This finding is consistent with research showing that cue and frame effects can depend heavily on how well the cue content matches the specific policy at hand [@feldmanClimateChangePolarizing2018], rather than operating as a uniform signal.

There was little support for H2. Liberal Democrats were largely unresponsive to both the Trump cue and the climate cue across the five actions, with only a marginally significant reduction in support for eliminating tax credits under the climate cue. This asymmetry between Republicans' and Democrats' responsiveness is consistent with prior work suggesting that out-group cues do not always produce the backlash that in-group cues produce for co-partisans [@merkleyPartyCuesNews2021]. It may also reflect a floor effect: because Democrats' opposition to actions such as coal leasing and tax credit elimination was likely already near the floor of the response scale in the control condition, there was less room for either cue to move their opinions further.

Finally, I find no evidence for H3 or H4. Respondents without a college degree were no more responsive to the Trump cue than those with a degree, and this null result held for both conservative Republicans and liberal Democrats. This suggests that, at least in the context of these energy actions, the purported strength of Trump's appeal among white working-class voters [@morganTrumpVotersWhite2018; @abramowitzUnitedStatesRacial2019] did not translate into greater education-based cue receptivity. This null finding is consistent with more recent work suggesting that cue receptivity is not simply a function of the cognitive resources associated with lower educational attainment [@bakkerUnderstandingPartisanCue2019]. It may also reflect the congruent nature of the Trump cue for conservative Republicans specifically. Prior work on cross-pressured cue-taking finds that education most clearly moderates responses to cues that conflict with a respondent's predispositions, with little role for education when a cue instead reinforces an existing partisan alignment [@vranceanuHowCrosspressuresAffect2022]. More broadly, this pattern is consistent with survey-experimental evidence that cues from high-profile partisan officials chiefly widen partisan polarization rather than persuade respondents to cross into out-group territory, regardless of respondents' education [@houstonHowEngagementHighProfile2026].

Although education did not moderate either cue's effect, it did independently condition support among the most ideologically committed respondents, in a two-way interaction with political beliefs. Conservative Republicans and liberal Democrats with a college degree were both less supportive than their non-college co-partisans of cancelling the nearly complete offshore wind project and of shortening the nuclear licensing timeline, and college-educated liberal Democrats were also less supportive of eliminating the wind and solar tax credits. This pattern runs opposite to the effect of a college degree among moderate and other respondents, for whom education is associated with greater, not lower, support for these same actions. This suggests that a college degree does not shift support for Trump's energy actions uniformly but instead interacts with political beliefs, dampening support among committed partisans on both sides while raising it among those without a strong partisan or ideological attachment.

For liberal Democrats, this pattern fits Ehret, Sparks, and Sherman's [-@ehretSupportEnvironmentalProtection2017] ideological-consistency model, in which more educated liberals hold more strongly pro-environmental positions than their less educated co-partisans. The pattern among conservative Republicans, however, runs opposite to what that model would predict, as more educated Republicans were less, not more, supportive of actions that curtail wind energy and nuclear expansion specifically. This suggests that education's role in amplifying partisan-consistent positions on energy may not be symmetric across parties, an asymmetry consistent with this study's broader finding that the climate and Trump cues themselves affected Republicans and Democrats differently. 

Together, these findings suggest that Trump's actions on energy are most likely to further polarize public opinion through their association with Trump himself, rather than through their connection to climate change. This is broadly consistent with the pattern of source-specific polarization found in the literature. The actions that showed the clearest Trump cue effects, coal leasing and nuclear licensing, involve energy sources where Republican support was already primed to move [@bergquistEnergyPolicyPublic2020], whereas the tax credit action, which touches wind and solar, showed no Trump cue effect at all. The results also suggest that Trump's influence on energy attitudes operates similarly to his documented effects on other issues, such as immigration and trade policy [@taylorTrumpEffectImmigration2026; @essigTrumpEffectPolitical2021], extending the "Trump effect" into the domain of energy policy specifically.

This study has some limitations. The survey was fielded during a specific and evolving period of the second Trump administration's energy actions, and respondents' reactions to the Trump cue may reflect the cumulative political context at that time rather than a more general and stable response to Trump. Additionally, the study captures respondents' immediate, single-exposure reactions to the cues, which may not reflect how cue effects accumulate or decay as respondents receive additional information about these actions over time.

Future research should examine whether the Trump cue effect documented here persists over time and whether it extends to other energy actions beyond the five examined here. Future research should also explore why the climate cue's effects were concentrated on the tax credit action specifically, as this may reveal more general principles about when framing an action's climate implications is more or less persuasive to conservative Republicans. Finally, future work should test whether the interaction between education and political beliefs replicates, as a pattern in which a college degree lowers support for these actions among partisans but raises it among non-partisans would not be well captured by either the low-information account of cue-taking or the diploma-divide literature.

Overall, this study provides experimental evidence that a cue identifying an energy action as coming from the Trump administration, rather than a cue linking the action to climate change, is the more consistent driver of polarized views on the second Trump administration's approach to energy. As the administration continues to reshape federal energy policy, these results suggest that public reactions to specific actions are likely to be structured more by attitudes toward Trump himself than by their connection to climate change, at least among conservative Republicans.

\newpage

# Appendix

```{r}
#| label: tbl-results-explore
#| tbl-cap: "Exploratory Model: Climate Cue x Education (Not Preregistered)"
results_table_explore
```

\newpage

# References