---
title: "Partisan vs Climate Cues and Public Opinion about Trump's Actions on Energy"
subtitle: "American Political Science Association Conference, September 2026"
title-slide-attributes:
  data-background-color: "#0064B1"
author:
  - name: Matthew C. Nowlin
    affiliation:
      - Associate Professor
      - Department of Political Science
      - University of Texas at Arlington
bibliography: /Users/matthewnowlin/Library/CloudStorage/OneDrive-UTArlington/01-RESEARCH/01-in-progress/cue-actions/references.bib
csl: /Users/matthewnowlin/Library/CloudStorage/OneDrive-UTArlington/01-RESEARCH/01-in-progress/cue-actions/american-political-science-association.csl
fig-cap-location: top
fontfamily: ebgaramond
link-bibliography: false
format:
  revealjs:
    theme: [pp.scss]
    logo: "UTAPoliticalScience.png"
    incremental: true
    include-in-header:
      text: |
        <style>
          #title-slide .title {
            font-size: 2em;
            color: white;
          }
          #title-slide .subtitle {
            font-style: italic;
          }
          .reveal:has(#title-slide.present) .slide-logo {
            display: none;
          }
        </style>
---




```{r setup, include=FALSE, message=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(".."))
```

```{r}
#| echo: false
#| include: false
#| warning: false
#| message: false

## loads data, builds the survey design, fits the H1-H4 models, and
## builds the tables/figures used below
source("scripts/manuscript-prep.R")
```


## Motivation

**_What shapes public support for President Trump's actions on energy_?**

* The second Trump administration has moved aggressively to promote fossil fuels and sideline renewable sources

* Historically, energy opinion was about cost and availability [@ansolabehereCheapCleanHow2016]

* Energy has become more polarized in recent years, likely through its connection to climate change

* Support for Trump's actions is likely polarized, but through **Trump**, or **climate change**?

## Trump's Actions on Energy

In the first year of his second term, the administration:

* Limited the eligibility of wind and solar projects for federal tax credits [@springTrumpAdministrationLimits2025]

* Required coal plants set to retire to remain open [@thakoreTrumpAdministrationOrders2025]

* Opened 13.1 million acres of federal land for coal leasing [@InteriorUnleashesAmerican2025]

* Cancelled a nearly complete offshore wind project [@wasserTrumpAdministrationPauses2025]

* Shortened the licensing timeline for new nuclear power plants [@usdepartmentofenergy9KeyTakeaways2025; @TrumpsLicensingGoals2025]

## Polarization on Energy Is Uneven

Partisan polarization differs across energy sources

* **Fossil fuels (coal)**: the largest partisan differences of any source [@hartPoliticalPolarizationSupport2021]

* **Wind and solar**: majorities across parties favor expansion; partisans disagree over *how much*, not *whether* [@bergquistEnergyPolicyPublic2020; @hamiltonChangeWindUS2018]

* **Nuclear**: somewhat more support on the right, but no sharp split [@yeoPartisanAmplificationRisk2014; @mcbethPoliticalIdeologyNuclear2023a]

* Republicans and Democrats also support renewables for *different reasons*; costs and energy security vs. climate change [@gustafsonRepublicansDemocratsDiffer2020]

## Cues and Cue-Taking

* Partisan cues are a heuristic: mass public infers their own position from the position of their party's elites [@kamWhoToesParty2005; @schaffnerPartisanHeuristicLowInformation2002]

* Best understood through motivated reasoning &mdash; processing information to protect prior partisan commitments [@taberMotivatedSkepticismEvaluation2006; @druckmanHowElitePartisan2013]

* On energy and climate, attaching a climate frame or a Democratic-source cue is often enough to decrease Republican support [@feldmanClimateChangePolarizing2018; @merkleyPartyCuesNews2021]

## A Trump Cue Is Distinct

* Individual elites, not parties in the abstract, tend to carry the most weight [@nicholsonPolarizingCues2012; @brulleShiftingPublicOpinion2012]

* A "Trump effect": Trump has moved opinion on immigration and trade [@taylorTrumpEffectImmigration2026; @essigTrumpEffectPolitical2021]

* His administration has taken concrete steps to expand fossil fuels and curb renewables [@gerrardTrumpsMultiProngedAttack2025], so the same dynamic may carry over to energy

* **Cues**: _a Trump cue activates expressive, cue-receptive partisanship; a climate cue taps the substantive climate concern that already divides the parties_

## Does Education Matter?

* Low-information account: less-educated citizens rely on party cues the most [@kamWhoToesParty2005; @merolaNumeracyPersuasiveEffect2016]

* But education can also enable *more effective* motivated reasoning [@kahanIdeologyMotivatedReasoning2013; @bakkerUnderstandingPartisanCue2019] and has *amplified* climate polarization among conservatives [@hamiltonEducationPoliticsOpinions2011]

* Trump's strength among non-college white voters &mdash; the "diploma divide" [@zingherTRENDSDiplomaDivide2022; @morganTrumpVotersWhite2018]

* **Education**: _if the Trump cue draws on identity-based appeal, its effects should concentrate among non-college partisans_

## Hypotheses

Partisan cue-taking and education

H1: _Conservative Republicans who receive the Trump or climate cue will be **more** supportive of Trump's actions than others_

H2: _Liberal Democrats who receive the Trump or climate cue will be **less** supportive than others_

H3: _Conservative Republicans **without a college degree** who receive the Trump cue will be more supportive than others_

H4: _Liberal Democrats **without a college degree** who receive the Trump cue will be less supportive than others_

::: {.notes}
Preregistered: https://osf.io/zkhqp/overview
:::


# Data and Analysis {.section-background}

## Survey

* Quota-based sample of US residents recruited through *Cloud Research*
  * Quotas from US Census data for age, gender, and race

* Internet-based, administered in QuestionPro

* Fielded April 10&ndash;15, 2026

* *N* = `r n_total`

* Iterative raking (`survey`, `anesrake`) to the 2023 ACS on age, race, gender, and education; Pew NPORS 2025 for party affiliation

## Experimental Design

Respondents randomly assigned to one of three conditions:

* **Trump cue** (*n* = `r n_trump`) &mdash; actions attributed to the *Trump Administration*

* **Climate cue** (*n* = `r n_climate`) &mdash; actions attributed to the *federal government*, plus a statement that scientists are concerned the actions could make climate change worse

* **Control** (*n* = `r n_control`) &mdash; actions attributed to the *federal government*, no mention of climate change

## Question Wording

Recently, the [*Trump Administration* / *Federal government*] took several actions to shape electricity generation in the U.S.

**Climate cue**: *Scientists are concerned that the actions associated with coal, wind, and solar energy will make climate change worse.*

Using a scale of 1 to 5, where 1 means strongly oppose and 5 means strongly support, how do you feel about the following actions of the [*Trump Administration* / *Federal government*]? *[items randomized]*

## Measures

* **Dependent variables**: support (1&ndash;5) for each of the five actions
  * Tax credits, coal open, coal leasing, cancel wind, nuclear licensing

* **Political beliefs**: conservative Republican, liberal Democrat
  * Moderate/other (moderate independents, cross-pressured partisans) is the excluded referent

* **College**: 1 = completed a college degree, 0 = did not


## Model

* Survey-weighted OLS (`svyglm`), one model per action

* Cue main effects, political beliefs, and college

* Interactions:
  * cue &times; political beliefs &rarr; H1, H2
  * Trump cue &times; beliefs &times; college &rarr; H3, H4

* Reference categories: control condition, moderate/other identity, no college degree


# Results {.section-background}

## Predicted Support by Cue and Political Beliefs

```{r}
#| echo: false
#| warning: false
#| fig-width: 11
#| fig-height: 4.5
fig_identity
```

## Predicted Support by Cue and Political Beliefs

**Conservative Republicans**

* Trump cue significantly *increased* support for opening federal lands to coal leasing and for shortening nuclear licensing
  * Increases for keeping coal plants open and cancelling the offshore wind project at *p* < 0.10

* Trump cue did **not** move support for eliminating wind and solar tax credits

* Climate cue *decreased* support for eliminating tax credits 

## Predicted Support by Cue and Political Beliefs

**Liberal Democrats**

* Largely unresponsive to both the Trump cue and the climate cue

* Reduction in support for eliminating tax credits under the climate cue (*p* < 0.10)

* Likely a floor effect as opposition to coal leasing and tax credit elimination was already near the bottom of the scale in the control condition

## Predicted Support by Trump Cue and College Degree

```{r}
#| echo: false
#| warning: false
#| fig-width: 11
#| fig-height: 6
fig_college
```

## Education Did Not Moderate the Cues

* No differences across the three-way interaction of Trump cue &times; college &times; political beliefs

* Non-college respondents were **no more** responsive to the Trump cue than college respondents &mdash; for both Republicans and Democrats (H3, H4 unsupported)

* But education *independently* conditions support among committed partisans:
  * College-educated conservative Republicans **and** liberal Democrats were *less* supportive of cancelling wind projects and shortening nuclear licensing than their non-college co-partisans

## Discussion

::: {.incremental}
* The **Trump cue**, not the climate cue, moved opinion &mdash; and only among conservative Republicans (H1 partly supported)

* No Trump cue effect on eliminating tax credits &mdash; the only action that *cuts* an existing renewable program rather than expanding or maintaining one

* The **climate cue** did little, except to *decrease* Republican support for eliminating tax credits

* **Liberal Democrats** were unresponsive to both cues (H2 unsupported)

* **Education** did not moderate cue effects (H3, H4 unsupported), but independently lowers support among committed partisans on both sides

* Trump's energy actions are more likely to polarize opinion through their association with **Trump himself** than through their connection to **climate change**
:::

# Thank You! {.section-background}

# Appendix {.section-background}

## Descriptive Statistics

```{r}
#| echo: false
desc_table
```

* Support for each action hovers near the midpoint of the 1&ndash;5 scale
  * Lowest: cancelling nearly complete wind projects (`r round(wtd_mean(d$fed.action.cancel.wind), 2)`)
  * Highest: shortening nuclear licensing (`r round(wtd_mean(d$fed.action.nuclear.licensing), 2)`)

## Difference in Means by Cue Condition

```{r}
#| echo: false
means_table
```

* The only significant raw difference: support for the nuclear regulatory changes is *lower* in the Trump condition than in the control

## Weighted OLS Models: Main Effects

```{r}
#| echo: false
modelsummary(
  models,
  coef_map = coef_map[c("trump.cue", "climate.cue", "conRep",
                        "libDem", "college", "(Intercept)")],
  gof_omit = gof_omit_pattern,
  stars = stars_map
) |> tinytable::style_tt(fontsize = 0.7)
```

::: {.notes}

Reference categories: control condition, moderate/other identity, no college degree

:::

## Weighted OLS Models: Interactions

```{r}
#| echo: false
modelsummary(
  models,
  coef_map = coef_map[c("trump.cue:conRep", "climate.cue:conRep",
                        "trump.cue:libDem", "climate.cue:libDem",
                        "trump.cue:college", "conRep:college", "libDem:college",
                        "trump.cue:conRep:college", "trump.cue:libDem:college")],
  gof_omit = gof_omit_pattern,
  stars = stars_map
) |> tinytable::style_tt(fontsize = 0.55)
```


::: {.notes}

`trump.cue` &times; identity tests H1/H2; `trump.cue` &times; identity &times; `college` tests H3/H4

:::


## References
